---
title: "Filters and Joins"
author: "ESellers"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# attached libraries
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
```

Get some fish.csv and kelp_fronds.xlsx data
**NOTE:** by default, R reads in the first excel sheet but you can specify which sheet you want to read in by defining either the sheet name or number

```{r read data}
fish <- read_csv(here("data/fish.csv")) # or here("data", "fish.csv")
kelp_abur <- read_excel(here("data/kelp_fronds.xlsx"))
# or add sheet = "sheetname or number" to specify nonfirst sheet
```

### Filter by an exact match

Exact match: "=="

Example 1: only keep observations from `fish` where the common_hname variable matches (==) "garibaldi"

```{r exact match filter}
# take fish and filter it where values in the common_name field exactly match garibaldi and name the resulting object fish_garibaldi
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
```

Activity: create a subset of fish called fish_over50, where the total count is greater than OR equal to 50

```{r activity filter fishover50}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
```

### Filter by this OR this

Option 1: Use the vertical line (pipe) `|` means "or"

```{r Filter by X OR y...}
fish_2sp <- fish %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith")
```

Option 2: filter for multiple matches within the same variable using the %in% operator (look for anything within a range of options)

Option 2: 

```{r Filter common_name for a range of ORd observations}
fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

Activity: subset from fish (fish_gar_2016) that keeps all observations where the year is 2016 OR the common_name is "garibaldi"

```{r Filter ORd values from diff cols}
fish_gar_2016 <- fish %>%
  filter(year == 2016 | common_name == "garibaldi")
```

### Filter to match this AND that using a comma `,` or `&`

```{r Filter ANDd values}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
```

To exclude things: for exclude where X "does not match" use `!=`

### stringr::str_Detect() for filtering on partial matches

```{r combo of filter and str_detect}
fish_black <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
```

ACtivity: What is the subset of data if common_name == "it"?

```{r partial match filter activity}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))
```

### `dplyr::*_join` functions

e.g. full_join(x,y, by = c("Species", "Year")) where the datasets x and y have these two cols in common

```{r merge two datasets that have common cols}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))
```

If colValue in Y matches colValue in X then merge

###left_join() to keep all from x, and merge matching observations from y with it (exclude things from y that don't have a match in x)
leftJoin (keep obs from dataframe Y ONLY if match to variable in dataframe X)

```{r leftJoin}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year","site"))
```

### inner_join() to merge data frames, only keeping observations with a match in both
Keep obs if there are complete matches across both data frames `inner_join()`

```{r innerJoin}
kelp_fish_innerjoin <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
# join where all observations contain a value in BOTH year AND site
```

### Pipe together multiple steps
Add an additional mutate step

- create an object called `my_fish_join` starting from `fish`
- filter to only include obs at site "abur" from the year 2017
- join kelp_abur data from to the resulting subset using left_join()
- add a new col using `mutate()` to calculate fish per kelp frond density (counts of fish / # of total fronds)

```{r}
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count/total_fronds)
```

Create a nice HTML table using kableExtra

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = 
                  "striped",
                full_width = FALSE)
```

ggplot(data = ?, aes(x = ?, y= ?)) + geom_?

```{r create a column graph}
ggplot(data = my_fish_join, aes(x = common_name, y = fish_per_frond)) + 
  geom_col(aes(fill = common_name))
```



































