---
title: "Pivot Lobsters"
author: "ESellers"
date: "1/27/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


Attach our packages and reading in data

```{r attach and install packages and read in data}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")
library(kableExtra) # install.packages("kableExtra")

# read in lobster data - and skip the first 4 rows because they contain metadata in this example
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"),
                        skip = 4)
```

Viewing our data with skimr 
- gives you a quick 'skim' or summary of the data like number of rows, any variables with missing values, etc.

```{r view data with skimr}
# these are equivalent
skimr::skim(lobsters) # explicitly stated from skimr
skim(lobsters) # not explicit
```

# Pivot Tables
%>% means "and then"
In R, Pivot tables are: `group_by() %>% summarize()`
The `n()` function will count observations or values in a field
```{r pivot table with 2 variables}
# group by year
lobsters %>% 
  group_by(year) %>% 
  summarize(count_by_year = n())

# group by year and site
lobsters %>% 
  group_by(year, site) %>% 
  summarize(count_by_year_site = n())

# group by site and year
lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n())
```

Summarize by multiple variables
The `na.rm= TRUE` instructs R to remove missing values from the mean function/calculation
And assign or label as an object called siteyear_summary
```{r pivot table with multiple variables}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm= TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

Table formatting with `kableExtra::kable()`
```{r table formatting with kableExtra}
siteyear_summary %>% 
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.  
**NOTE:** The addition of the r at the beginning tells RMarkdown that you want this code snippet to run when you knit this.

## Activity
1. build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm` variable anhd a function called `median()`

```{r}
siteyear_median_lobster_size <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            median_size_mm = median(size_mm, na.rm= TRUE))
```


2. create a ggplot and save. For example:
```{r}
ggplot(data = siteyear_median_lobster_size, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))
```


3. sync this to github


