---
title: "Creating graphs with ggplot2"
author: "ESellers"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse. 

We will install these packages:

- tidyberse
- readxl: to read in excel files (in many ways)
- here: to help with filepaths

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")
```

# California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
# can also be written (in this case but good practice to include the 'here' package that knows the full file path:
# ca_np <- read_csv(here("data", "ca_np.csv"))
# also be written:
# ca_np <- read_csv("data/ca_np.csv")

ci_np <- read_excel(here("data","ci_np.xlsx"))
```

### Make my own first graph!
We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:

```{r using ggplot with ci_np}
# ggplot (dataframe, aesthetic(variables)) +creates a soft return
# geom_line(no specifics) line graph/plot type
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point() #select different types of charts line, column, area etc.
```

Save and name a base line of code as an object for CI (Channel Islands) plots:
```{r create and name the base line of code for Channel Islands plots}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point() # now use the newly named object to insert the reusable code snippet
```

### Now let's start customizing!

- Line / point colors with: color
- Change line type with: linetype
- Change point style with: pch
- Change transparency with: apha

```{r plot CI visitation again}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Activity: create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch) of a non-default size (size = #), and that are an updated color.

```{r Activity create a ggplot scatterplot}
gg_base +
  geom_point(
    color = "dodgerblue",
    pch = 2,
    size = 4
  )
```

### Update graph aesthetics based on variables
Size and color of the points will change based on the number of visitors
The blue color in the example below is the RStudio default - not something we defined.
```{r Change point size and color based on variable values}
gg_base +
  geom_point(
    aes(size = visitors,
    color = year),
    pch = 10
  )

# Add an argument to update the color based on the year value
```


