---
title: "Tidying Session Day2"
author: "ESellers"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE) # echo=true is to show code in the Rmarkdown doc

# Attaching packages. Note remember to alwasy do Ctrl Enter for each of these to 'run' them.
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```

### Read in inverts.xlsx

```{r read in data}
inverts <- read_excel(here("data", "inverts.xlsx")) # or "data/inverts.xlsx" (NOTE R uses the forward slash)
```

### `pivot_longer()`
We can use a single colon (:) to indicate a series of sequential columns (e.g. A:D)
```{r pivot_longer convert variables split across columns}
inverts_long <- pivot_longer(data = inverts, 
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count") #put the values for each year in a new column sp_count
                             
# What if you want some in series and some NOT in series?
# c('2016':'2018', '2020')
```

What if we want to get it back into wide format?

```{r pivot_wider}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name, 
              values_from = sp_count)
```

### Use `janitor::clean_names()` (package::functionInthePackage) to tidy up field labels

```{r inverts_wide with clean colNames}
inverts_wide <- inverts_wide %>% 
  clean_names()

# clean_names default case is lowercase_snake_case
```

### `tidyr::unite` to combine info from multiple columns
Example:merge the site code and year into a single column called "site_year"

```{r combine multiple cols}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year", 
        c(site, year),
        sep = "_") #col = use quotes whenever creating a new variable name; sep = separator delimiter
```


### Activity: Create a new ojbect called 'inverts_moyr' starting from inverts_long that united the month and year cols into a SINGLE col named "mo_yr usig a slash "/" as the sepearator... then try some other fun separators

```{r Activity using unite function}
inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr",
        c(year, month),
        sep ="_")
```

### `tidyr::separate()` to separate a single col value into multiple cols

```{r Separate col into multi-cols}
inverts_sep <- inverts_unite %>% 
  separate(col = site_year, into = c("my_site", "my_year"))
```

### `stringr` + `mutate()` functions

Example: using built-in `tree` dataset with mutate() to add 5 to all of the tree height values

```{r mutate to add a new col}
trees_new <- trees %>% 
  mutate(height_update = Height + 5) #if you use the same col_name R will overwrite the orig col
```

### `str_replace()` incombo with `mutate` to change string pattern

```{r use str_replace inside mutate function to change a string pattern}
ca_abbr <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name, 
                       pattern = "california",
                       replacement = "CA"))
```

Example: In the common_name column, you want to replace the word 'spiny' with 'awesome'

```{r}
awesome_lobsters <- inverts %>% 
  mutate(common_name =
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "awesome"))
```

```{r TEST nested stringr and mutate}
# this didn't quite work - keeps adding badly named cols but does replace all 3 values
nested_str_replace <- inverts %>% 
  mutate(common_name =
           str_replace(common_name, pattern = "california", replacement = "CA") %>%mutate(common_name =
           str_replace(common_name, pattern = "spiny", replacement = "awesome") %>% 
  mutate(common_name =
           str_replace(common_name, pattern = "snail", replacement = "mollusc")) %>% 
```



























